### 2017-04-01 manuscript/publication-abstract.tex
Final read through and submission (almost)

\begin{abstract}
  Many of the challenges facing interpretation and knowledge synthesis from life cycle assessment (LCA) studies stem from the inability of study authors and readers to formally agree on the structure and content of life cycle inventory models.  This article presents a framework for LCA study authors to publish inventory models in a structured format based on a mathematical formulation of a \emph{foreground study}.  Using linked semantic data as a foundation, the publication framework offers a mechanism to provide both machine readability and unambiguous human interpretation of published studies.  Implementing the framework requires only that the LCA community provides a way for study authors to make common reference to shared data sources, including background databases, elementary flows, and impact assessment methods, that are widely used in many studies.  The framework contributes to ongoing efforts within industrial ecology to improve the reproducibility and verifiability of scholarly works, and if implemented, plots a course toward distributed, platform-independent computation and validation of LCA results.
\end{abstract}

>>>>

\begin{abstract}
  Many of the challenges facing %interpretation and
  knowledge synthesis from life cycle assessment (LCA) studies stem from the inability of study authors and readers to formally agree on the structure and content of life cycle inventory models.  This article presents a framework for LCA study authors to publish inventory models in a structured format based on a mathematical formulation of a \emph{foreground study}.  Using linked semantic data as a foundation, the publication framework %offers a mechanism to provide
  provides for both machine readability and unambiguous human interpretation of published studies.  To implement the framework, the LCA community should find ways to enable study authors and data users to make common reference to shared data sources, including background databases, elementary flows, and impact assessment methods, that are widely used in many studies.  The framework contributes to ongoing efforts within industrial ecology to improve the reproducibility and verifiability of scholarly works, and if implemented, plots a course toward distributed, platform-independent computation and validation of LCA results.
\end{abstract}

### 2017-04-01 manuscript/publication-components.tex
Final read through and submission (almost)

The collection of foreground nodes and their reference flows expresses a de facto definition of the scope of the model, which is enclosed by an implicit system boundary.  %Most published studies disclose at least the structure of their foreground models by providing a graphical system diagram in the LCA report.  These ``box-and-arrow'' diagrams can be interpreted as graphs and used to produce an adjacency matrix that shows the locations of non-zero elements of Af.  However, the model disclosed in the graph may be a simplified form of the actual model, and the adjacency matrix itself is not sufficient to reproduce the matrix unless the exchange values themselves are also explicitly reported.

>>>>

The foreground is enclosed by an implicit system boundary which includes all foreground nodes and their reference flows.  Therefore, the list of foreground nodes and reference flows can act as a de facto definition of the effective scope of the model.  %Most published studies disclose at least the structure of their foreground models by providing a graphical system diagram in the LCA report.  These ``box-and-arrow'' diagrams can be interpreted as graphs and used to produce an adjacency matrix that shows the locations of non-zero elements of Af.  However, the model disclosed in the graph may be a simplified form of the actual model, and the adjacency matrix itself is not sufficient to reproduce the matrix unless the exchange values themselves are also explicitly reported.

### 2017-04-01 manuscript/publication-conclusion.tex
Final read through and submission (almost)

Before the goal of easily reproducible foreground models can be realized, the linked data foundation must be laid.   The four entity classes found in foreground studies, namely foreground flows, background processes, elementary flows, and characterization quantities, must shift from being recorded locally to being stored on the Semantic Web, where their shared meaning can be curated.  Fortunately, this is easy to do: authoritative sources simply need to publish hyperlinks to their content. Providers of aggregated inventory data should make available static hyperlinks to their content, so that licensees and data users can make shared reference to the same entities, even if the data may not be accessible to the general public. %Thinkstep already does this: every process inventory they publish has a stable link on the Web\footnote{\href{http://gabi-documentation-2016.gabi-software.com/xml-data/processes/f8dddbcb-f08c-4217-99de-0b59d2c11bb1.xml}{http://gabi-documentation-2016.gabi-software.com/xml-data/processes/[UUID].xml} where [UUID] is replaced with the process's unique identifier. For example, \texttt{f8dddbcb-f08c-4217-99de-0b59d2c11bb1} corresponds to production of 1,3-Butadiene, via catalytic dehydrogenation of butane, in Germany in 2015.}.
Similarly, impact assessment providers should publish their characterization data in a way that allows for automatic query and validation.  Both of these tasks can be accomplished independently through the efforts of members of the LCA community; however, the accuracy and validity of the data would be improved if the work were done by the data providers themselves.

>>>>

Before the goal of easily reproducible foreground models can be realized, the linked data foundation must be laid.   The four entity classes found in foreground studies, namely foreground flows, background processes, elementary flows, and characterization quantities, must shift from being recorded locally to being stored on the Semantic Web, where their shared meaning can be curated.  Fortunately, this is easy to do: authoritative sources simply need to maintain static hyperlinks to their content, so that licensees and data users can make shared reference to the same entities, even if the data may not be accessible to the general public. %Thinkstep already does this: every process inventory they publish has a stable link on the Web\footnote{\href{http://gabi-documentation-2016.gabi-software.com/xml-data/processes/f8dddbcb-f08c-4217-99de-0b59d2c11bb1.xml}{http://gabi-documentation-2016.gabi-software.com/xml-data/processes/[UUID].xml} where [UUID] is replaced with the process's unique identifier. For example, \texttt{f8dddbcb-f08c-4217-99de-0b59d2c11bb1} corresponds to production of 1,3-Butadiene, via catalytic dehydrogenation of butane, in Germany in 2015.}.
Similarly, impact assessment providers should publish their characterization data in a way that allows for automatic query and validation.  Both of these tasks can be accomplished independently through the efforts of members of the LCA community; however, the accuracy and validity of the data would be improved if the work were done by the data providers themselves.  A shared semantic framework for foreground flows would also support collaborative development of inventory data resources.

### 2017-04-01 manuscript/publication-discussion.tex
Final read through and submission (almost)

An LCA foreground study can be published concisely by distinguishing foreground components from background components, explicitly identifying all entities involved in the study, and then reporting the data values that make up the submatrices in Eq.~\ref{eqn:study}.  As the examples demonstrate, a structured publication can be created that allows a reader to critically review, validate, and accurately reproduce the product system model without requiring additional information.  The entity map provides the foundation for a linked-data publication by explicitly naming all externally-referenced entities involved in the foreground model.  Ultimately it should be possible to name entities using linked data identifiers or URIs, allowing unambiguous, automatic retrieval of metadata about the entities.

The input matrices ($A_f$, $A_d$, or $B_f$) or aggregation results ($\tilde{\mathbf{x}}$, $\tilde{\mathbf{a}}_d$, $\tilde{\mathbf{b}}_f$), together with the entity map, form a research object that can be easily generated and shared with colleagues.  By separating foreground and background, the research object satisfies the objective for attribution, because only the information included in the research object is attributable to the study author; whereas information that came from the background database is excluded from the publication.

This framework allows an author a great deal of flexibility in publishing their model.  Any of the components or intermediate results visualized in Figure~\ref{fig:foreground} may be selected for publication at any level of detail.  By publishing a set of linked data identifiers corresponding to the rows and entries of the matrices, the author ensures that the data user will be able to identify the foreground and determine the background processes used. Again because the background is excluded, the size of the research object is kept down, and licensing limitations associated with the distribution of proprietary data are avoided.

\subsection{Reproducing LCIA Results}

@@

For background scores, the results cannot be validated without access to the background database.  Again, the characterization factors used would need to be disclosed along with the background LCI results.  Because background databases tend to be updated infrequently, it is important for a background dependency reference to include information about which version was used.  Assuming datasets within a particular background database do not change, it would be possible to create an archive of unit impact scores for each background process and LCIA method supported by the database.  These values could then be independently validated.

In the context of structured publication, a reader of a structured publication could obtain the unit LCIA scores for each background dataset referenced, and use them to validate the background LCIA result. If the background database is subject to licensing restrictions, it may be necessary to require readers or data users to independently obtain background LCIA scores.  However, it may also be possible for the author to publish unit impact scores for background processes used in the model without violating licensing terms. In this case, it is much easier to reproduce published results, though the reader may still wish to check the author's published unit scores for accuracy.  If unit impact scores for background processes are included in the publication, then it is a simple matter for a reader to perform a contribution analysis and inspect the sensitivity of the results to foreground parameters.  This is exemplified in the ``organic potato'' publication provided in the supplementary materials.

@@

>>>>

An LCA foreground study can be published concisely by distinguishing foreground components from background components, explicitly identifying all entities involved in the study, and then reporting the data values that make up the submatrices in Eq.~\ref{eqn:study}.  As the examples demonstrate, a structured publication can be created that allows a reader to critically review, validate, and accurately reproduce the product system model without any additional information.  The entity map provides the foundation for a linked-data publication by explicitly naming all externally-referenced entities involved in the foreground model.  Ultimately it should be possible to name entities using linked data identifiers or URIs, allowing unambiguous, automatic retrieval of metadata about the entities.

The input matrices ($A_f$, $A_d$, and $B_f$) or aggregation results ($\tilde{\mathbf{x}}$, $\tilde{\mathbf{a}}_d$, $\tilde{\mathbf{b}}_f$), together with the entity map, form a research object that can be easily generated and shared with colleagues.  By separating foreground and background, the research object satisfies the objective for attribution, because only the information attributable to the study author is included in the research object; whereas information that came from the background database is excluded from the publication.

This framework allows an author a great deal of flexibility in publishing their model.  Any of the components or intermediate results visualized in Figure~\ref{fig:foreground} may be selected for publication at any level of detail.  By publishing a set of linked data identifiers corresponding to the rows and entries of the matrices, the author ensures that the data user will be able to identify the foreground and determine the background processes used. Again because the background is excluded, the size of the research object is kept down, and licensing limitations associated with the distribution of proprietary data are avoided. Use of the research object to selectively exclude private data is discussed in the supplementary materials.

\subsection{Validating LCIA Results}

@@

For background scores, the results cannot be validated without access to the background database.  Again, the characterization factors used would need to be disclosed along with the background LCI results.  Because background databases are occasionally updated, it is important for a background dependency reference to include information about which version was used.  Assuming datasets within a particular background database do not change, it would be possible to create an archive of unit impact scores for each background process and LCIA method supported by the database.  These values could then be independently shared and validated.

In the context of structured publication, a reader of the publication could obtain the unit LCIA scores for each background dataset referenced, and use them to validate the background LCIA result. If the background database is subject to licensing restrictions, it may be necessary to require readers or data users to independently obtain background LCIA scores.  However, it may also be possible for the author to publish unit impact scores for background processes used in the model without violating licensing terms. In this case, it is much easier to reproduce published results, though the reader may still wish to check the author's published unit scores for accuracy.  If unit impact scores for background processes are included in the publication, then it is a simple matter for a reader to perform a contribution analysis and inspect the sensitivity of the results to foreground parameters.  This is exemplified in the ``organic potato'' publication provided in the supplementary materials.

@@

%%%% TODO TODO TODO TODO!!!!!!! save it for the revision
\subsection{Interlinking Foreground Models}

Current practice does not provide a way for study authors to easily provide their models to collaborators, or for  independently created models to be linked together.  However, structured publication of foreground studies could allow independent authors to work together on complex models.  Large foregrounds can be easily partitioned into fragments simply by selectively including or excluding columns of $A_f$, $A_d$, and $B_f$ in different publications, allowing for independent maintenance of different subsystems. Similarly, an isolated product system described in a structured format could easily be combined with other models of complementary systems to form a larger system.  Independent foreground systems that both reference the same shared cutoff flows could be automatically linked without any intervention by the authors.

In both cases, the interlinking of foreground models rests on the principle of shared reference to common entities.  Linked semantic data provides an enabling technology, both allowing standalone foreground publications to be automatically linked to data resources and to one another, and by providing interpretive services such as dataset discovery, flow matching, unit conversion, and metadata retrieval.  This represents an opportunity for further research and development.

### 2017-04-01 manuscript/publication-example.tex
Final read through and submission (almost)

The product system models in most LCAs can be described as foreground studies (Eq.~\ref{eqn:foreground}).  In fact, LCI databases contain product systems that can also be modeled as foreground studies because they describe products that are not required by the background.  In this section, two product systems selected from LCI databases are used to illustrate the concept of structured publication.  The systems selected because are complex enough to illustrate the premise but simple enough to review easily.

Each system is illustrated as a table that shows the foreground model, cutoff flows, background dependencies and foreground emissions included in the system.  The table omits numeric data from most of the table for clarity. Instead, a black square indicates the presence of a nonzero value.  Aggregation results $\tilde{\mathbf{x}}$, $\tilde{\mathbf{a}}_d$, and $\tilde{\mathbf{b}}_f$ are also reported.  The systems are also reproduced as structured publications in the supplementary materials, formatted as Excel spreadsheets.

@@

Ecoinvent version 3 is provided in three different system models that reflect different linking strategies.  The example is drawn from the ``Allocation: Point of substitution'' or APOS model, which includes 11,420 processes that produce 12,966 product flows.  Of these, 10,282 are background flows and the rest are foreground flows.

The example system reports production of organic potatoes supplied to the global market (Table \ref{table:potato}).  The foreground includes nine nodes, of which six (nodes 3 through 8) form a strongly connected component associated with the production of potato seeds.  The Ecoinvent database's use of ``markets'' as mixer processes is evident in the table: node 0 is a mixer process that combines Swiss (``CH'') production (node 1) weighted at 2.4\%  with rest-of-World (``RoW'') production (node 2) weighted at 97.6\%.  That same market split (2.4 / 97.6) can also be seen in nodes 7 and 8.

@@

The structured publication of the organic potato model reports the $A_f$, $A_d$, and $B_f$ matrices in full rather than sparse format to enable a more accessible visual review.  In addition, the $E$ matrix is reported in full, for all emissions found in the model, rather than only the emissions found in the foreground.  This enables the publication to be used to visually evaluate different implementations of the same impact assessment method in the context of the inventory model (Ecoinvent vs ILCD implementations of ReCiPe Midpoint marine eutrophication; see supplementary materials for details).

>>>>

LCI databases often contain product systems that can be modeled as foreground studies because they describe products that are not required by the background.  In this section, two product systems selected from LCI databases are used to illustrate the concept of structured publication.  The systems selected because are complex enough to illustrate the premise but simple enough to review easily.

Each system is illustrated as a table that shows the foreground model, cutoff flows, background dependencies and foreground emissions included in the system.  Aggregation results $\tilde{\mathbf{x}}$, $\tilde{\mathbf{a}}_d$, and $\tilde{\mathbf{b}}_f$ are also reported.  The table omits numeric data from most of the table for clarity. Instead, a black square indicates the presence of a nonzero value.  The systems are also reproduced as structured publications in the supplementary materials, formatted as Excel spreadsheets.

@@

Ecoinvent version 3 is provided in three different system models that reflect different linking strategies.  The example is drawn from the ``Allocation at the point of substitution'' or APOS model, which includes 11,420 processes that produce 12,966 product flows.  Of these, 10,282 are background flows and the rest are foreground flows.

The example system reports production of organic potatoes supplied to the global market (Table \ref{table:potato}).  The foreground includes nine nodes, of which six (nodes 3 through 8) form a cyclic dependency associated with the production of potato seeds.  The Ecoinvent database's use of ``markets'' as mixer processes is evident in the table: node 0 is a mixer process that combines Swiss (``CH'') production (node 1) weighted at 2.4\%  with rest-of-World (``RoW'') production (node 2) weighted at 97.6\%.  That same market split (2.4 / 97.6) can also be seen in nodes 7 and 8.

@@

The structured publication of the organic potato model reports the $A_f$, $A_d$, and $B_f$ matrices in full rather than sparse format to enable a more visually immediate review.  In addition, the $E$ matrix is reported in full, for all emissions found in the model, rather than only the emissions found in the foreground.  This enables the publication to be used to visually evaluate different implementations of the same impact assessment method in the context of the inventory model (Ecoinvent vs ILCD implementations of ReCiPe Midpoint marine eutrophication; see supplementary materials for details).

### 2017-04-01 manuscript/publication-foreground.tex
Final read through and submission (almost)

\subsection{Model fragments and foreground inputs and outputs}

Conventionally, the foreground flows are entirely inside the system boundary: they are produced by one foreground node and consumed by one or more others.  In the conventional case, the only product flow which crosses the implicit system boundary is the reference flow, which provides the functional unit of the study given in $\tilde{\mathbf{y}}$.  By implication, the difference $\tilde{\mathbf{x}} - \tilde{\mathbf{y}}$ represents the magnitudes of interior flows within the foreground.

However, these flows may not all necessarily be interior.  Most complex life cycle models involve nested model designs, in which low-level model components are encapsulated within higher-level models.  A foreground matrix that describes an incomplete life cycle model could be called a model ``fragment.''  In model fragments, some foreground flows represent product flows consumed within the fragment but not produced there, or vice versa. These flows represent inputs and outputs to the foreground model that are not further exchanged among the foregrond nodes, but neither are they terminated to a background process or emitted into the environment.  Thus input/output flows can be detected by identifying foreground flows that show up as row entries but have no column entries.

The input/output flows provide a mechanism to link separate foreground fragments together.  Two foreground models that both specify foreground flows with the same identity can be linked together simply by combining their entries in a common set of $A_f$, $A_d$, and $B_f$ matrices.  Similarly, large foregrounds can be easily partitioned into fragments simply by selectively including or excluding columns of $A_f$, $A_d$, and $B_f$ in different publications.
In isolation from other foreground models, input/output flows may be considered to be ``cutoff flows'' that must be terminated by a data user.

@@

\subsection{Foreground Dependencies and Foreground Emissions}

@@

A structured publication could include any of these elements, bundled into a research object that meets the objectives laid out above.  This research object must include:
\begin{itemize}
\item \emph{tabular data} which enables the data user to construct the matrices and perform the computation;
\item an \emph{entity map}, which allows for precise interpretation of the elements of the study.
\end{itemize}
These two components are discussed below.

@@

The sparse representation assumes the matrix is largely zeros, and only reports the non-zero elements.  A sparse representation could also take the form of a delimited file, except each line would contain exactly three entries: the row index, column index, and value of the non-zero matrix entry.  Given that the matrices being represented are largely made up of zeros, the sparse approach is preferable to the full one.  The exception is $B_x$ or any column thereof, which is likely to contain no non-zero entries.  Publication of $B_x$ is a special case that will be discussed in a later section.

@@

The entity map must provide some minimal information about the entities in each of the four groups discussed above:
\begin{itemize}
\item \textbf{Foreground flows}: Because the foreground is specific to the model, the publication itself \emph{is} the basis for the semantic meaning of the foreground flows. Authors should simpy provide enough detail that the significance of each flow is clear to a data user.  Each foreground flow must specify its \emph{reference unit}, which is the amount of the flow that corresponds to a unit activity of the foreground node.  The data values reported elsewhere in the column must be normalized to this unit.
    % and does not necessarily relate to any other publication, there are no restrictions on how the individual flows should be named. T
    %It is also not necessary for the foreground flows to be mapped to anything outside the model. Rather,
\item \textbf{Background processes}: A properly identified background data set should correspond to a distinct aggregated unit process dataset from a background data provider.  The reference units of the background processes should already be part of the background data publication and do not need to be reproduced in the foreground publication except for clarity.  
\item \textbf{Elementary flows}: The best practices for identifying elementary flows are a matter of ongoing discussion and research.  At minimum, each elementary flow should be identified in terms of (1) its chemical composition, using a reference nomenclature, CAS number, chemical formula, or other concrete identifier; and (2) the environmental compartment with which it is being exchanged.  Geographic locale may also be a part of an elementary flow specification.  However, if the flow originates in the foreground, the locale may be inferred from the locale of the originating foreground node.
  \item \textbf{Characterization quantities}: The LCIA characterization method is already required to be unambiguously reported in current LCA practice.  If a structured publication includes results, it should also include a link to an authoritative reference for the method(s) computed.  
\end{itemize}

The entity map should include all of the foreground flows, background processes, elementary flows, and characterization quantities referred to in the publication.  Taken together, the entity map provides a working formal definition of the scope of the model.  Even in the absence of the numerical data that make up the input matrices or aggregation results, this \emph{functional scope} can be used as a basis to describe the contents of the model and formally state the system boundary of the foreground.

>>>>

\subsubsection{Model fragments and foreground inputs and outputs}

Conventionally, the foreground flows are entirely inside the system boundary: they are produced by one foreground node and consumed by one or more others.  In this case, the only product flow which crosses the implicit system boundary is the reference flow, which provides the functional unit of the study given in $\tilde{\mathbf{y}}$.  By implication, the difference $\tilde{\mathbf{x}} - \tilde{\mathbf{y}}$ represents the magnitudes of interior flows within the foreground.

However, not all these flows are necessarily interior.  Most complex life cycle models involve nested model designs, in which low-level model components are encapsulated within higher-level models.  A foreground matrix that describes an incomplete life cycle model could be called a model ``fragment.''  In model fragments, some foreground flows represent product flows consumed within the fragment but not produced there, or vice versa. These flows represent inputs and outputs to the foreground model that may be produced or consumed by other foreground systems.
%are not further exchanged among the foregrond nodes, but neither are they terminated to a background process or emitted into the environment.  Thus
Input/output flows can be detected by identifying foreground flows that show up as row entries but have no column entries in the $A_f$, $A_d$, or $B_f$ matrices.  If there is no other foreground system that connects to an input/output flow, it may be considered to be a ``cutoff flow'' that is excluded from the model scope.
%
%%The input/output flows provide a mechanism to link separate foreground fragments together.  Two foreground models that both specify foreground flows with the same identity can be linked together simply by combining their entries in a common set of $A_f$, $A_d$, and $B_f$ matrices.  Similarly,
%%In isolation from other foreground models, input/output flows may be considered to be ``cutoff flows'' that must be terminated by a data user.

@@

\subsubsection{Foreground Dependencies and Foreground Emissions}

@@

A structured publication could include any of these elements, bundled into a research object that meets the objectives laid out above.  This research object must include:
\begin{itemize}
\item \emph{tabular data} which enables the data user to construct the matrices and perform the computation;
\item an \emph{entity map}, which allows for precise interpretation of the elements of the study.
\end{itemize}
%These two components are discussed below.

@@

The sparse representation assumes the matrix is largely zeros, and only reports the non-zero elements.  A sparse representation could also take the form of a delimited file, except each line would contain exactly three entries: the row index, column index, and value of the non-zero matrix entry.  Given that the matrices being represented are largely made up of zeros, the sparse approach is preferable to the full one.  %The exception is $B_x$ or any column thereof, which is likely to contain no non-zero entries.  Publication of $B_x$ is a special case that will be discussed in a later section.

@@

The entity map must provide some minimal information about the entities in each of the four groups discussed above:
\begin{itemize}
\item \textbf{Foreground flows}: Because the foreground is specific to the model, the publication itself \emph{is} the basis for the semantic meaning of the foreground flows. Authors should simpy provide enough detail that the significance of each flow is clear to a data user.  Each foreground flow must specify its \emph{reference unit}, which is the amount of the flow that corresponds to a unit activity of the foreground node.  The data values reported elsewhere in the column must be normalized to this unit.
    % and does not necessarily relate to any other publication, there are no restrictions on how the individual flows should be named. T
    %It is also not necessary for the foreground flows to be mapped to anything outside the model. Rather,
\item \textbf{Background processes}: A properly identified background data set should correspond to a distinct aggregated unit process dataset from a background data provider.  The reference units of the background processes should already be part of the background data publication and do not need to be reproduced in the foreground publication except for clarity.  
\item \textbf{Cutoff flows}: Cutoff flows are explicit exclusions from the system boundary.  They may represent inputs or outputs from the foreground that are produced by another foreground system, or they may represent dependencies for which no background process is available.  Thus, they are semantically similar to foreground flows. They may or may not make reference to externally defined entities.  However, complex foreground models can be published in parts by making shared reference to a set of cutoff flows.
\item \textbf{Elementary flows}: The best practices for identifying elementary flows are a matter of ongoing discussion and research.  At minimum, each elementary flow should be identified in terms of (1) its chemical composition, using a reference nomenclature, CAS number, chemical formula, or other concrete identifier; and (2) the environmental compartment with which it is being exchanged.  Geographic locale may also be a part of an elementary flow specification.  However, if the flow originates in the foreground, the locale may be inferred from the locale of the originating foreground node.
  \item \textbf{Characterization quantities}: The LCIA characterization method is already required to be unambiguously reported in current LCA practice.  If a structured publication includes results, it should also include a link to an authoritative reference for the method(s) computed.  
\end{itemize}

The entity map should include all of the foreground flows, background processes, cutoff flows, elementary flows, and characterization quantities referred to in the publication.  Taken together, the entities in the map provide a working formal definition of the scope of the model.  Even in the absence of the numerical data that make up the input matrices or aggregation results, this \emph{functional scope} can be used as a basis to describe the contents of the model and formally state the system boundary of the foreground.

### 2017-04-01 manuscript/publication-intro.tex
Final read through and submission (almost)

Life cycle assessment (LCA) is a standard methodology to estimate the potential environmental impacts of products or services by modeling the network of industrial processes that must occur to deliver a quantitative measure of utility to a consumer.  The technique is well-established, widely practiced, and draws on an extensive body of standardization and scholarship.
%Performing an LCA first requires preparing an inventory of the amounts of environmental interventions associated with the product, such as resources extracted or emissions released.  The inventory is the input to an impact assessment computation, in which the inventory items are scaled by their relative significance according to impact categories of interest, and summed together.  %% 55 words
Increasingly, the results of LCA studies are appearing in environmental declarations, corporate sustainability reports, and product marketing information.  LCA is also gaining prominence as a tool for developing and evaluating environmental policy.  
%
At the core of an LCA is a model of a ``product system,'' which represents a collection of ``processes $\ldots$ performing one or more defined functions'', that describes the life cycle of a product \citep{iso14044}.  Product system models encode information about how products are manufactured, distributed, used by the consumer, and what happens to them after they are disposed.  A product system can be divided into a foreground, which denotes the portions of the life cycle whose operations are directly modeled by the study, and a background, which represents the global industrial system \citep{SETAC_inventory_1998}.  While the foreground is often modeled from direct observation or other primary data, background processes are typically drawn from a life cycle inventory (LCI) database prepared by a third party.
%A relatively small number of comprehensive LCI databases exist, owing to their vast data requirements and the complexity of preparing and maintaining them \citep{UNEP_2011}.  The product system model itself is often constructed in specialized software, which is used to compute and analyze impact assessment results.  $$ 45 words

Although there has been considerable effort to normalize and harmonize %%important methodological decisions around %% -4 +1 
LCI database design methodology and data interchange \citep{UNEP_2011, JRC_ILCD_ELCD_2013, Mila_e_Canals_2015, Ingwersen_JLCA_2015}, less attention has been paid to harmonizing product system models.  These models are often highly complex and include countless modeling decisions, approximations, and assumptions made by a study author \citep{Lloyd2007, reap2008_I}.  Consequently, while comparative evaluations can be made regarding alternative cases or scenarios within a single study, comparisons across studies are much more challenging \citep{Heath2012, Henriksson2014}.  Even when relatively simple and homogeneous systems are considered, wide variations can be found in the results of studies from different authors \citep{van_der_Harst_2013, Turconi_2013}.
%Some types of product systems may require coordination on key methodological issues before results can be compared \citep{Ahlgren_2015}, while others face more fundamental challenges that arise from the complexity of the supply chain and the lack of suitable inventory or proxy data \citep{Bull2014}. 43 words
The lack of transparency in reporting,
particularly regarding definitions of study scope and system boundary,
is a crucial factor challenging the interpretation of results in comparative analysis \citep{Cleary2009, Laurent_2014}.
%In particular, this lack of transparency may manifest as a imprecision in the definitions of study scope and system boundary \citep{Laurent_2014, Turconi_2013}. %% -23 + 9
One consequence
%% of the wide variation in model construction, as well as the lack of consistent transparency in reporting, %% -18
is the high cost and uneven rigor of critical review, in which the complexities of LCA come face to face with the limitations of current practice \citep{Curran2014}.

As LCA gains prominence, particularly in the policy realm, these problems become more acute.  If LCA is to be used in policy, it will necessarily have the effect of recommending certain product systems, technologies, or approaches at the expense of others; yet in that event it is critically important that there be a consensus among stakeholders that the quantitative results are credible and well-supported \citep{Rainville_2015, McManus_2015}.  As noted above, this level of consensus is hard to realize, in part because of the conflicting implications inherent in different analytic modes.
%It is well known that allocation decisions can have dispositive effects on the outcome of a study; yet no consensus exists for specifying allocation methods (e.g. \cite{Hanes_2015, Pelletier_2014}), much less evaluating their significance. %% 33 words
Other factors can also render study findings unreliable or highly contingent.  Plevin et al.\ famously observed in the strongest terms that careless reporting of attributional LCA study results can distort the significance of findings in a policy context \citep{Plevin_2013}, a shortcoming that can be found in other modes as well \citep{Brandao_2014}.  In the scope of Environmental Product Declarations (EPDs), studies within an industry are supposed to be rendered comparable by adhering to a common product category rule (PCR) \citep{Fet_2006}.  However, this prescription is insufficient to ensure the comparability of different declarations, even if they use the same PCR \citep{Modahl_2012}.  Moreover, PCRs are often themselves not unique and can include wide variations in scope and boundary requirements \citep{Subramanian_2012}, leading to ambiguity.  %%PCRs themselves tend, like EPDs and other LCA studies, to be written qualitatively, to be unstructured, and to be open to considerable interpretation

>>>>

Life cycle assessment (LCA) is a standard methodology to estimate the potential environmental impacts of products or services by modeling the network of industrial processes that must occur to deliver %a quantitative measure of utility
them to a consumer.  The technique is well-established, widely practiced, and draws on an extensive body of standardization and scholarship.
%Performing an LCA first requires preparing an inventory of the amounts of environmental interventions associated with the product, such as resources extracted or emissions released.  The inventory is the input to an impact assessment computation, in which the inventory items are scaled by their relative significance according to impact categories of interest, and summed together.  %% 55 words
Increasingly, the results of LCA studies are appearing in environmental declarations, corporate sustainability reports, and product marketing information.  LCA is also gaining prominence as a tool for developing and evaluating environmental policy.  
%
At the core of an LCA study is a model of a ``product system,'' which represents a collection of ``processes $\ldots$ performing one or more defined functions'', that describes the life cycle of a product \citep{iso14044}.  Product system models encode information about how products are manufactured, distributed, used by the consumer, and what happens to them after they are disposed.  A product system can be divided into a foreground, which denotes the portions of the life cycle whose operations are directly modeled by the study, and a background, which represents the global industrial system \citep{SETAC_inventory_1998}.  While the foreground is often modeled from direct observation or other primary data, background processes are typically drawn from a life cycle inventory (LCI) database prepared by a third party.
%A relatively small number of comprehensive LCI databases exist, owing to their vast data requirements and the complexity of preparing and maintaining them \citep{UNEP_2011}.  The product system model itself is often constructed in specialized software, which is used to compute and analyze impact assessment results.  $$ 45 words

Although there has been considerable effort to normalize and harmonize %%important methodological decisions around %% -4 +1 
LCI database design methodology and data interchange \citep{UNEP_2011, JRC_ILCD_ELCD_2013, Mila_e_Canals_2015, Ingwersen_JLCA_2015}, less attention has been paid to harmonizing product system models.  These models are often highly complex and include countless modeling decisions, approximations, and assumptions made by a study author \citep{Lloyd2007, reap2008_I}.  Consequently, while comparative evaluations can be made regarding alternative cases or scenarios within a single study, comparisons across studies are much more challenging \citep{Heath2012, Henriksson2014}.  Even when relatively simple and homogeneous systems are considered, wide variations can be found in the results of studies from different authors \citep{van_der_Harst_2013, Turconi_2013}.
%Some types of product systems may require coordination on key methodological issues before results can be compared \citep{Ahlgren_2015}, while others face more fundamental challenges that arise from the complexity of the supply chain and the lack of suitable inventory or proxy data \citep{Bull2014}. The  ## 43 words
A lack of transparency in reporting,
particularly regarding definitions of study scope and system boundary,
is a crucial challenge to the interpretation of results in comparative analysis \citep{Cleary2009, Laurent_2014}.
%In particular, this lack of transparency may manifest as a imprecision in the definitions of study scope and system boundary \citep{Laurent_2014, Turconi_2013}. %% -23 + 9
One consequence
%% of the wide variation in model construction, as well as the lack of consistent transparency in reporting, %% -18
is the high cost and uneven rigor of critical review, in which the complexities of LCA come face to face with the limitations of current practice \citep{Curran2014}.

As LCA gains prominence, particularly in the policy realm, these problems become more acute.  If LCA is to be used in policy, it will necessarily have the effect of recommending certain product systems, technologies, or approaches at the expense of others; yet in that event it is critically important that there be a consensus among stakeholders that the quantitative results are credible and well-supported \citep{Rainville_2015, McManus_2015}.  As noted above, this level of consensus is hard to realize, in part because of the conflicting implications inherent in different analytic modes.
%It is well known that allocation decisions can have dispositive effects on the outcome of a study; yet no consensus exists for specifying allocation methods (e.g. \cite{Hanes_2015, Pelletier_2014}), much less evaluating their significance. %% 33 words
Other factors such as allocation strategy \citep{Pelletier_2014} can also render study findings unreliable or highly contingent.  Plevin et al.\ famously observed in the strongest terms that careless reporting of attributional LCA study results can distort the significance of findings in a policy context \citep{Plevin_2013}, a shortcoming that can be found in other modes as well \citep{Brandao_2014}.  In the scope of Environmental Product Declarations (EPDs), studies within an industry are supposed to be rendered comparable by adhering to a common product category rule (PCR) \citep{Fet_2006}.  However, this prescription is insufficient to ensure the comparability of different declarations, even if they use the same PCR \citep{Modahl_2012}.  Moreover, PCRs are often themselves not unique and can include wide variations in scope and boundary requirements \citep{Subramanian_2012}, leading to ambiguity.  %%PCRs themselves tend, like EPDs and other LCA studies, to be written qualitatively, to be unstructured, and to be open to considerable interpretation
